StateSpaceModels.jl: a Julia Package for \\ Time-Series Analysis in a State-Space Framework

### 2019-07-07 paper/paper.tex
updating paper

The filtering step derives the predictive state as well as its covariance matrix at every time period via the Kalman filter, i.e., it estimates
%
\begin{align}
    a_{t+1} &= \mathbb{E}[\alpha_{t+1} | Y_{t}], \\
    P_{t+1} &= \mathbb{V}[\alpha_{t+1} | Y_{t}],
\end{align}
%
where $Y_{t}$ denotes the set of time periods $y_{1}, \dots, y_{t}$.

@@

Consider the following simulated time series composed by a growing trend plus a Gaussian noise. Additionally, we remove observations 10 through 20 to treat them as missing:
%
\begin{lstlisting}[language = Julia]
# Create growing trend series with Gaussian noise
y = collect(1:0.25:20) + 0.5*randn(77)
# Remove observations 10 through 20
y[10:20] .= NaN
# Specify the state-space model and estimate it
model = linear_trend(y)
ss = statespace(model)
\end{lstlisting}
%
The results of the estimation step are presented in Fig. \ref{fig:missing_observations}. The filtered state and smoothed state estimates can be accessed in \texttt{ss.filter.a} and \texttt{ss.smoother.alpha}, respectively. It is clear in Fig. \ref{fig:missing_observations} that the filter and smoother are able to effectively capture the growing linear trend of the series even in the period where observations are missing.
%
\begin{figure}[h]
	\centering
	\includegraphics[width=\columnwidth]{Figures/missing_obs.pdf}
	\caption{Automatic completion of missing observations in the filtering and smoothing.}
	\label{fig:missing_observations}
\end{figure}

@@

\begin{itemize}
    \item ~\texttt{model}: the previously defined \texttt{StateSpaceModel}.
    \item ~\texttt{filter}: a \texttt{FilteredState} structure representing the result of the Kalman filter. It contains the predictive state and the innovations as well as their covariance matrices at each time period. Furthermore, there is a flag indicating if steady state was attained and the time period when it was attained.
    \item ~\texttt{smoother}: a \texttt{SmoothedState} structure representing the result of the smoothing procedure. It contains the smoothed state as well as its covariance matrix at each time period.
    \item ~\texttt{covariance}: a \texttt{StateSpaceCovariance} structure which contains the covariance matrices of the observation and state noises.
    \item ~\texttt{filter\_type}: the defined filter type.
    \item ~\texttt{optimization\_method}: the defined optimization method.
\end{itemize}

>>>>

The filtering step derives the predictive and filtered states as well as their covariance matrices at every time period via the Kalman filter. The predictive state and its variance are given by
%
\begin{align}
    a_{t+1} &= \mathbb{E}[\alpha_{t+1} | Y_{t}], \\
    P_{t+1} &= \mathbb{V}[\alpha_{t+1} | Y_{t}],
\end{align}
%
while the filtered state and its variance are given by
%
\begin{align}
    a_{t|t} &= \mathbb{E}[\alpha_{t} | Y_{t}], \\
    P_{t|t} &= \mathbb{V}[\alpha_{t} | Y_{t}],
\end{align}
%
where $Y_{t}$ denotes the set of time periods $y_{1}, \dots, y_{t}$.

@@

Consider the following simulated time series composed by a growing trend plus a Gaussian noise. Additionally, we remove observations 10 through 20 to treat them as missing:
%
\begin{lstlisting}[language = Julia]
# Create growing trend series with Gaussian noise
y = collect(1:0.25:20) + 0.5*randn(77)
# Remove observations 10 through 20
y[10:20] .= NaN
# Specify the state-space model and estimate it
model = linear_trend(y)
ss = statespace(model)
\end{lstlisting}
%
The results of the estimation step are presented in Fig. \ref{fig:missing_observations}. The predictive, filtered, and smoothed state estimates can be accessed in \texttt{ss.filter.a}, \texttt{ss.filter.att}, and \texttt{ss.smoother.alpha}, \mbox{respectively}. It is clear in Fig. \ref{fig:missing_observations} that the filter and smoother are able to effectively capture the growing linear trend of the series even in the period where observations are missing.
%
\begin{figure}[h]
	\centering
	\includegraphics[width=\columnwidth]{Figures/missing_obs.pdf}
	\caption{Automatic completion of missing observations in the filtering and smoothing.}
	\label{fig:missing_observations}
\end{figure}

@@

\begin{itemize}
    \item ~\texttt{model}: the previously defined \texttt{StateSpaceModel}.
    \item ~\texttt{filter}: a \texttt{FilterOutput} structure representing the result of the Kalman filter. It contains the predictive and filtered states as well as the innovations and their covariance matrices at each time period. Furthermore, there is a flag indicating if steady state was attained and the time period when it was attained.
    \item ~\texttt{smoother}: a \texttt{SmoothedState} structure representing the result of the smoothing procedure. It contains the smoothed state as well as its covariance matrix at each time period.
    \item ~\texttt{covariance}: a \texttt{StateSpaceCovariance} structure which contains the covariance matrices of the observation and state noises.
    \item ~\texttt{filter\_type}: the defined filter type.
    \item ~\texttt{optimization\_method}: the defined optimization method.
\end{itemize}
